A comparative study of batch construction strategies \\ for recurrent neural networks in MXNet

### 2017-04-19 paper/paper.tex
final pass

  We extracted 16-dimensional MFCC vectors as in \cite{menne16:chime4System} from the six sub-corpora
  and used them as features for the neural network models. The context-dependent phonemes extracted from the lexicon were clustered into 
  1500 state labels using a classification and regression tree (CART).
  We trained a three state HMM without skip transitions applied it in a forced alignment mode to the training data in order to obtain a frame-wise labeling. A three layer 
  bi-directional LSTM followed a softmax layer was trained to minimize the cross-entropy of the softmax output and the CART label sequence at each frame. Optimization was done with the Adam optimizer and a constant learning rate of 0.01. We used MXNet \cite{mxnet} for experiments using the bucketing approach and RETURNN \cite{doetsch2017:returnn} for the other methods.

  After training, state posteriors were extracted from the softmax for each observation of the evaluation set. The posteriors were then decoded in the RWTH 
  speech recognition toolkit RASR \cite{rybach2011:rasr} using the hybrid HMM approach. A 4-gram LM 
  was used during decoding with a language model scale of 12.

  \section{Experiments} \label{sec:experiments}
   In order to provide some perspective on the required context on the CHiME-4 task, 
   we performed experiments on sub-utterance (chunk) which are presented in 
   Table \ref{tab:chime:chunk}. In the first columns we evaluate different sub-utterance lengths, while measuring the utterances that were processed per second, the memory required and the word error rate (WER) on the evaluation set of the CHiME-4 database. We hereby constrained batches to only 
   contain 5,000 frames in total, such that the same number of updates is constant in all experiments. We can observe that 
   while large speed-ups can be obtained when training is done in this fashion, full-utterance context is
   required for optimal performance. However, it is worth noting that the memory requirement decreases 
   significantly when sub-utterance training is applied and especially for unusually long utterances, 
   sub-utterance training might be the only way to fit the data into GPU memory.

   For training on full sequences we conducted experiments with different batch construction strategies.
   The results are reported in Table \ref{tab:chime:batch}, where the first three rows show results for trivial sequence ordering 
   methods and the last two rows provide a direct comparison of the bucketing approach as 
   it is implemented in MXNet and the alternated sorting approach as proposed in this paper.

	\begin{table}[tbp]
		\centering
		\caption{Training time and recognition performance when training is done 
		on sub-utterances level. The first column shows the maximal sub-utterance length 
		after partitioning of the original sequence. The last line shows the results 
		if no partitioning into sub-utterances is performed. }
		\label{tab:chime:chunk}
		\begin{tabular}{lrrr}
			\hline
			Chunk size              & Utt./sec & Memory [GB] & WER [\%] \\
			\hline
			10     					&	36.7   & 1.6         &	21.3     \\
			50  					&	31.1   & 1.6		 &  10.1     \\
			100 					& 	29.6   & 1.6		 &	9.2      \\
			500						& 	17.3   & 1.6         &	9.0      \\
			$\max$					&	7.0	   & 5.4         &	8.9      \\
			\hline
		\end{tabular}
	\end{table}

	\begin{table}[tbp]
		\centering
		\caption{An evaluation of different sequence batch construction methods on the CHiME-4 database. Training time per epoch, memory consumption are presented in the first two columns, while the last columns shows the word error rate of the corresponding system
		when decoded with the RASR speech recognition toolkit.}
		\label{tab:chime:batch}
		\begin{tabular}{lrrr}
			\hline
			Approach                & Utt./sec & Memory [GB] & WER [\%]          \\
			\hline
			Random 					& $\sim$ 7.0     & $\sim$ 5.4		 & $\sim$ 8.9		 \\
			Sorted					& 10.2    &	4.2 & 10.2		 \\
			\hline
			Bucketing (MXNet)		& 9.5     &	6.3		 & 9.6		 \\
			Proposed (8 bins)	    & 10.1	   & 4.8     & 9.9		 \\
			Proposed (64 bins)	    & 10.0     & 5.3     & 9.5		 \\
			Proposed (256 bins)	    & 8.8	   & 6.0     & 9.1		 \\
			\hline
		\end{tabular}
	\end{table}

@@

  \section{Conclusions}
    In this work we presented a novel strategy to construct sequence-level batches for recurrent 
    neural network acoustic model training. While not much attention is given to the topic 
    of batch construction, we demonstrate that different strategies can lead to large variations 
    both in training time and recognition performance. Most deep-learning frameworks rely on a bucketing approach by clustering sequences of similar length into bins and to create batches from each bin individually. We showed that we can achieve a better runtime performance using a simpler batch design, by partitioning a shuffled sequence order and to sort the partitions in alternating order. The method was evaluated on the ChiME-4 noisy speech recognition task and 
    compared to standard approaches like random sequence shuffling and the bucketing approach of MXNet,
    where our method was able to reach a better trade-off between training time and recognition performance while being easier to parametrize than the bucketing method.

>>>>

  We extracted 16-dimensional MFCC vectors as in \cite{menne16:chime4System} from the six sub-corpora
  and used them as features for the neural network models. The context-dependent HMM states were clustered into 
  1500 classes using a classification and regression tree (CART).
  We trained a GMM-based baseline model with three state HMM without skip transitions in order to obtain a frame-wise labeling of the training data. A network of three 
  bi-directional LSTM layers followed by a softmax layer was trained to minimize the frame-wise cross-entropy. Optimization was done with the Adam optimizer and a constant learning rate of 0.01. We used MXNet \cite{mxnet} for experiments using the bucketing approach and RETURNN \cite{doetsch2017:returnn} for the other methods.

  After training, the state posterior estimates from the neural network are normalized by the state priors and used as likelihoods in a conventional hybrid HMM decoder using the RASR toolkit~\cite{rybach2011:rasr}.
  A 4-gram LM 
  was used during decoding with a language model scale of 12.

  \section{Experiments} \label{sec:experiments}
   In order to provide some perspective on the impact of temporal context on the CHiME-4 task, 
   we performed experiments on sub-utterance (chunk) which are presented in 
   Table~\ref{tab:chime:chunk}. For different sub-utterance lengths, we report the processing speed measured in utterances per second, the memory required and the word error rate (WER) on the evaluation set of the CHiME-4 database. Here we constrained batches to only 
   contain 5,000 frames in total, such that the overall number of updates is constant in all experiments. We can observe that 
   while large speed-ups can be obtained when training is done in this fashion, full-utterance context is
   required for optimal performance. However, it is worth noting that the memory requirement decreases 
   significantly when sub-utterance training is applied. In particular, for unusually long utterances, 
   sub-utterance training might be the only way to fit the data into GPU memory.

   For training on full sequences we conducted experiments with different batch construction strategies.
   The results are reported in Table \ref{tab:chime:batch}, where the first two rows show results for trivial sequence ordering 
   methods and the last rows provide a direct comparison of the bucketing approach as 
   it is implemented in MXNet and the alternated sorting approach as proposed in this paper.

	\begin{table}[tbp]
		\centering
		\caption{Training time and recognition performance when training is done 
		on sub-utterance level. The first column shows the maximal sub-utterance length 
		after partitioning of the original sequence. The last row shows the results 
		obtained without partitioning into sub-utterances. }
		\label{tab:chime:chunk}
		\begin{tabular}{lrrr}
			\hline
			Chunk size              & Utt./sec & Memory [GB] & WER [\%] \\
			\hline
			10     					&	36.7   & 1.6         &	21.3     \\
			50  					&	31.1   & 1.6		 &  10.1     \\
			100 					& 	29.6   & 1.6		 &	9.2      \\
			500						& 	17.3   & 1.6         &	9.0      \\
			$\max$					&	7.0	   & 5.4         &	8.9      \\
			\hline
		\end{tabular}
	\end{table}

	\begin{table}[tbp]
		\centering
		\caption{An evaluation of different sequence batch construction methods on the CHiME-4 database. Training time per epoch, memory consumption are presented in the first two columns, while the last column shows the word error rate of the corresponding acoustic model on the evaluation set.}
		\label{tab:chime:batch}
		\begin{tabular}{l@{} rrr}
			\hline
			Approach                & Utt./sec & Memory [GB] & WER [\%]          \\
			\hline
			Random 					& $\sim$ 7.0     & $\sim$ 5.4		 & $\sim$ 8.9		 \\
			Sorted					& 10.2    &	4.2 & 10.2		 \\
			\hline
			Bucketing (MXNet)		& 9.5     &	6.3		 & 9.6		 \\
			Proposed (8 bins)	    & 10.1	   & 4.8     & 9.9		 \\
			Proposed (64 bins)	    & 10.0     & 5.3     & 9.5		 \\
			Proposed (256 bins)	    & 8.8	   & 6.0     & 9.1		 \\
			\hline
		\end{tabular}
	\end{table}

@@

  \section{Conclusions}
    In this work we presented a novel strategy to construct sequence-level batches for recurrent 
    neural network acoustic model training. While not much attention is given to the topic 
    of batch construction, we demonstrate that different strategies can lead to large variations 
    both in training time and recognition performance. Most deep-learning frameworks rely on a bucketing approach by clustering sequences of similar length into bins and to draw batches from each bin individually. We showed that we can achieve a better runtime performance using a simpler batch design, by partitioning a shuffled sequence order and to sort the partitions in an alternating order. The method was evaluated on the ChiME-4 noisy speech recognition task and 
    compared to standard approaches like random sequence shuffling and the bucketing approach of MXNet,
    where our method was able to reach a better trade-off between training time and recognition performance while being easier to parametrize than the bucketing method.
